---
title: "average statistics"
author: "sToney"
date: "2023-05-10"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
theme_set(theme_minimal())
```


```{r}
extreme_per_stage <- read_csv(here::here("data/drought_index.csv")) %>% 
  mutate(type = factor(type, levels = c("single","early","late"))) 


get_wilcoxon_test_res <- function(var_name){
  early_or_late <- extreme_per_stage %>% 
    filter(type != "single") %>% 
    group_split(type)
  data1 <- early_or_late[[1]] %>% 
    select(station,type,year,stage,all_of(var_name)) %>% 
    pivot_wider(names_from = "stage", values_from = var_name)
  data2 <- early_or_late[[2]] %>% 
    select(station,type,year,stage,all_of(var_name)) %>% 
    pivot_wider(names_from = "stage", values_from = var_name)
  map_dbl(1:10,\(col_num){
    test_data <- 
      inner_join(
        data1 %>% 
          rename(value_1 = col_num+3 ) %>% 
          select(station, year, value_1),
        data2 %>% 
          rename(value_2 = col_num+3 ) %>% 
          select(station, year, value_2),
        by = join_by(station, year)
      )
    wilcox.test(
      test_data$value_1,
      test_data$value_2,
      paired = TRUE
    ) %>% 
      pluck("p.value")
  })
}




get_mw_test_res <- function(var_name) {
  # Mann–Whitney U Test
  base_split <- extreme_per_stage %>% 
    select(station, type, year, stage, all_of(var_name)) %>% 
    group_split(type) %>% 
    map(pivot_wider,names_from = "stage", values_from = var_name)
  
  p1 <- map_dbl(1:10,\(i){
    wilcox.test(
      pull(base_split[[1]], c(str_c("period_",1:9),"overall")[[i]]),
      pull(base_split[[2]], c(str_c("period_",1:9),"overall")[[i]]),
      paired = FALSE
    ) %>% 
      pluck("p.value")
  })
  p2 <- map_dbl(1:10,\(i){ wilcox.test(
    pull(base_split[[1]], c(str_c("period_",1:9),"overall")[[i]]),
    pull(base_split[[3]], c(str_c("period_",1:9),"overall")[[i]]),
    paired = FALSE
  ) %>% 
      pluck("p.value")
  })
  list(early = p1, late = p2)
}

wilcoxon_test_res <- 
  map_dfc(c("freq","intns","prop"), get_wilcoxon_test_res) %>% 
  set_names(c("freq","intns","prop")) %>% 
  mutate(stage = c(str_c("period_",1:9),"overall"))
mw_test_res <- map(c("freq","intns","prop"),get_mw_test_res)
```





```{r}
generate_sig_table <- function(wilcoxon_test_res, mw_test_res) {
  sig_table_1 <- wilcoxon_test_res %>% 
    pivot_longer(freq:prop, names_to = "index", values_to = "p") %>% 
    mutate(group1 = "early", group2 = "late") %>% 
    # mutate(value = round(value,2)) %>% 
    mutate(p_star = case_when(
      p <= 0.001 ~ "***",
      p <= 0.01 ~ "**",
      p <= 0.05 ~ "*",
      TRUE ~ ""
    )) %>% 
    mutate(p = str_c(scales::label_scientific(1)(p), p_star)) %>% 
    mutate(y.position = case_when(
      index == "freq" ~ 0.038,
      index == "intns" ~ 2.56,
      TRUE ~ 0.0045
    ))
  
  sig_table_2 <- map(mw_test_res,pluck,"early") %>% 
    reduce(bind_cols) %>% 
    set_names(c("freq","intns","prop")) %>% 
    mutate(group1 = "single", group2 = "early") %>% 
    mutate(stage = c(str_c("period_",1:9),"overall")) %>% 
    pivot_longer(freq:prop, names_to = "index",values_to = "p") %>% 
    mutate(p_star = case_when(
      p <= 0.001 ~ "***",
      p <= 0.01 ~ "**",
      p <= 0.05 ~ "*",
      TRUE ~ ""
    )) %>% 
    mutate(p = str_c(scales::label_scientific(1)(p), p_star)) %>% 
    mutate(y.position = case_when(
      index == "freq" ~ 0.043,
      index == "intns" ~ 2.85,
      TRUE ~ 0.0051
    ))
  
  
  sig_table_3 <- map(mw_test_res,pluck,"late") %>% 
    reduce(bind_cols) %>% 
    set_names(c("freq","intns","prop")) %>% 
    mutate(group1 = "single", group2 = "late") %>% 
    mutate(stage = c(str_c("period_",1:9),"overall")) %>% 
    pivot_longer(freq:prop, names_to = "index",values_to = "p") %>% 
    mutate(p_star = case_when(
      p <= 0.001 ~ "***",
      p <= 0.01 ~ "**",
      p <= 0.05 ~ "*",
      TRUE ~ ""
    )) %>% 
    mutate(p = str_c(scales::label_scientific(1)(p), p_star)) %>% 
    mutate(y.position = case_when(
      index == "freq" ~ 0.048,
      index == "intns" ~ 3.14,
      TRUE ~ 0.0057
    ))
  return(
    list(
      sig_table_1 = sig_table_1,
      sig_table_2 = sig_table_2,
      sig_table_3 = sig_table_3
    )
  )
}
generate_sig_table(wilcoxon_test_res,mw_test_res) %>% attach()
```

```{r}
```{r}
index_dict <- c("freq" = "Frequency","intns" = "Intensity", "prop" = "Proportion")
extend_limits <- tribble(
  ~index, ~ value, ~ type,
  "freq", 0.053, "single",
  "intns", 3.35, "early",
  "prop", 0.008, "late"
)

label_tibble <- 
  crossing(
    index = factor(c("freq","intns","prop") , levels = c("freq","intns","prop") ),
    stage = factor(c("overall",paste0("period_",1:9)),levels = c("overall",paste0("period_",1:9)))
  ) %>% 
  mutate(label = paste0("(",1:30,")"))
panel_names <- c("Frequency","Intensity","Proportion","Overall",
                 paste0("Period ",1:9)) %>% 
  set_names(c('freq','intns','prop',"overall",paste0("period_",1:9)))
panel_names_ch <- c("频率","强度","占比","整体",
                    paste0("阶段",1:9)) %>% 
  set_names(c('freq','intns','prop',"overall",paste0("period_",1:9)))
boxplot_summarised_group <- extreme_per_stage %>% 
  pivot_longer(freq:prop, names_to = "index") %>% 
  group_by(type, index, stage) 
boxplot_data <- boxplot_summarised_group %>% 
  group_keys() %>% 
  bind_cols(
    boxplot_summarised_group %>% 
      group_split() %>% 
      map_dfc(
        ~boxplot.stats(.$value, do.conf = FALSE, do.out = FALSE)$stats
        ) %>% 
      t() %>% 
      as_tibble() %>% 
      rename(lower = V1, pctl_25 = V2, pctl_50 = V3, pctl_75 = V4, upper = V5)) 

```
```


```{r}
violin_plot_data <-  extreme_per_stage %>% 
  pivot_longer(freq:prop, names_to = "index") %>% 
  left_join(boxplot_data, by = c("type","index","stage")) %>% 
  filter(value < upper)
sig_table_11 <- sig_table_1 %>% 
  mutate(y.position = if_else(index == "prop",sqrt(y.position)+0.003,y.position))
sig_table_22 <- sig_table_2 %>% 
  mutate(y.position = if_else(index == "prop",sqrt(y.position)+0.006,y.position))
sig_table_33 <- sig_table_3 %>% 
  mutate(y.position = if_else(index == "prop",sqrt(y.position)+0.009,y.position))



box_violin_plot <- function(index_type = "freq") {
  boxplot_data %>% 
    filter(index == index_type) %>% 
    ggplot()+
    geom_errorbar(aes(type, ymin = lower, ymax = upper), width = 0.8)+
    geom_boxplot(
      aes(type, middle = pctl_50,
          upper = pctl_75,lower = pctl_25,
          ymax = pctl_75, ymin = pctl_25,
          fill = type),
      stat = "identity")+
    ggpubr::stat_pvalue_manual(
      filter(sig_table_11,index == index_type), size = 7,
      tip.length = ifelse(index_type == "prop",0.01,0.03))+
    ggpubr::stat_pvalue_manual(
      filter(sig_table_22,index == index_type),size = 7,
      tip.length = ifelse(index_type == "prop",0.01,0.03))+
    ggpubr::stat_pvalue_manual(
      filter(sig_table_33,index == index_type),size = 7,
      tip.length = ifelse(index_type == "prop",0.01,0.03))+
    geom_text(
      aes(x = -Inf, y = Inf, label = label),
      data = filter(label_tibble,index == index_type),
      size = 6,
      hjust = -0.2,
      vjust = 1.3,
      inherit.aes = FALSE)+
    facet_wrap(vars(stage), strip.position = "bottom",
               nrow = 1,labeller = as_labeller(panel_names))+
    geom_point(data = filter(extend_limits,index == index_type), 
               aes(type, value),color = NA)+
    ggthemes::scale_fill_tableau(labels = paste0(c("Early","Single","Late"),
                                                 " Rice"))+
    scale_x_discrete(NULL, labels = NULL)+
    labs(x = "", y = panel_names[index_type], fill = "")+
    theme_bw()+
    {if (index_type != "prop") theme(strip.text = element_blank())}+
    theme(axis.ticks.x = element_blank(),
          legend.position = "bottom",
          axis.title.y.left = element_text(angle = 0,vjust = 0.5,size = 30),
          strip.background =element_rect(fill="white",color = "white"),
          text = element_text(family = "serif",size = 40))
}
p1 <- box_violin_plot("freq")+
  scale_y_continuous(expand = c(0,0), labels = scales::label_percent())
p2 <- box_violin_plot("intns")+
  scale_y_continuous(expand = c(0,0))
p3 <- box_violin_plot("prop")+
  scale_y_continuous(expand = c(0,0), 
                     labels = scales::label_percent(),
                     trans = scales::sqrt_trans())

p <- p1/p2/p3+
  plot_layout(guides = "collect") &
  theme(legend.position='bottom',
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,-10,-10),
        legend.text = element_text(size = 35),
        legend.key.size = unit(1, 'cm'))

ggsave(here::here("plot/01EDA_comaprison_en.jpg"),width = 10,height = 6.5, dpi = 300)

```

```{r}
sig_table_111 <- sig_table_11 %>% 
  select(-p) %>% rename(p = p_star)
sig_table_222 <- sig_table_22 %>% 
  select(-p) %>% rename(p = p_star)
sig_table_333 <- sig_table_33 %>% 
  select(-p) %>% rename(p = p_star)
box_violin_plot <- function(index_type = "freq") {
  boxplot_data %>% 
    filter(index == index_type) %>% 
    ggplot()+
    # geom_errorbar(aes(type, ymin = ave*0.9,ymax= ave+std))+
    geom_errorbar(aes(type, ymin = lower, ymax = upper), width = 0.8)+
    # gghalves::geom_half_boxplot(
    geom_boxplot(
      aes(type, middle = pctl_50,
          upper = pctl_75,lower = pctl_25,
          ymax = pctl_75, ymin = pctl_25,
          # ymax = upper, ymin = lower,
          fill = type),
      stat = "identity")+
    ggpubr::stat_pvalue_manual(
      filter(sig_table_111,index == index_type),size = 15,
                               tip.length = ifelse(index_type == "prop",0.01,0.03))+
    ggpubr::stat_pvalue_manual(
      filter(sig_table_222,index == index_type),size = 15,
                               tip.length = ifelse(index_type == "prop",0.01,0.03))+
    ggpubr::stat_pvalue_manual(
      filter(sig_table_333,index == index_type),size = 15,
                               tip.length = ifelse(index_type == "prop",0.01,0.03))+
    geom_text(aes(x = -Inf, y = Inf, label = label),
              data = filter(label_tibble,index == index_type),
              size = 7,
              hjust = -0.2,
              vjust = 1.3,
              inherit.aes = FALSE)+
    facet_wrap(vars(stage), strip.position = "bottom",
               nrow = 1,labeller = as_labeller(panel_names_ch))+
    geom_point(data = filter(extend_limits,index == index_type), aes(type, value),color = NA)+
    ggthemes::scale_fill_tableau(labels = paste0(c("早","中","晚"),"稻"))+
    scale_x_discrete(NULL, labels = NULL)+
    labs(x = "", y = panel_names_ch[index_type], fill = "")+
    theme_bw()+
    {if (index_type != "prop") theme(strip.text = element_blank())}+
    theme(axis.ticks.x = element_blank(),
          axis.text.y.left  = element_text(family = "serif",size=30),
          axis.text.x  = element_text(family = "serif",size=30),
          # legend.position = "bottom",
          axis.title.y.left = element_text(angle = 0,vjust = 0.5),
          strip.background =element_rect(fill="white",color = "white"),
          text = element_text(family = "SyHei",size=40))
}
p1 <- box_violin_plot("freq")+
  scale_y_continuous(expand = c(0,0), labels = scales::label_percent())
p2 <- box_violin_plot("intns")+
  scale_y_continuous(expand = c(0,0))
p3 <- box_violin_plot("prop")+
  scale_y_continuous(expand = c(0,0), 
                     labels = scales::label_percent(),
                     trans = scales::sqrt_trans())
library(patchwork)
library(showtext)
p <- p1/p2/p3+
  plot_layout(guides = "collect") &
  theme(legend.position='bottom',text = element_text(family = "SyHei"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,-10,-10),
        legend.text = element_text(size = 35),
        legend.key.size = unit(1, 'cm'))

font_add("SyHei", "C:\\Users\\sToney\\AppData\\Local\\Microsoft\\Windows\\Fonts\\SourceHanSerifSC-Regular.otf")
showtext_auto()
ggsave(here::here("plot/01EDA_comaprison_ch.png"),
       plot = p,
       width = 10,height = 6, dpi = 300,
       device=ragg::agg_png)
```

